Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-09-09 multilevel_metaregression/manuscript_with_multilevel_metan.qmd
some changes for revision

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Raphaelle Delpech
    email: raphaelle.delpech.18@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Glyn Lewis
    email: glyn.lewis@ucl.ac.uk
    affiliations:
        - ref: ucl_1        
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: custom-reference-doc.docx
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    numer-sections: true
editor: visual
bibliography: "emotion_concepts_paper.bib"
csl: "vancouver.csl"
font:
  family: "Calibri"
prefer-html: true
fig-format: png
fig-cap-location: top
---

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

@@

# For referencing
bib_data <- ReadBib(here("emotion_concepts_paper.bib"))

```{r global_options, include=FALSE}
knitr::opts_chunk$set(tab.cap.fp_text = officer::fp_text_lite(italic = FALSE, bold = TRUE))
```

@@

For adolescent depression, there are limited head-to-head trials of medication and psychotherapy, and hence recommendations are derived from indirect comparisons of treatment efficacy. The National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend psychotherapy (specifically cognitive behaviour therapy (CBT) or interpersonal therapy (IPT)) over medication in most cases [@niceDepressionChildrenYoung2019]. This is in keeping with two sources of evidence relating to child and adolescent depression: meta-analyses of medication randomised controlled trials (RCTs) that cast doubt on the efficacy of antidepressants, with the exception of fluoxetine [@ciprianiComparativeEfficacyTolerability2016]; and meta-analyses of psychotherapy RCTs that conclude psychotherapy to be efficacious [@cuijpersEffectsPsychologicalTreatments2021]. However, a recent network meta-analysis (NMA) [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments using both direct and indirect (i.e. treatment A with treatment C, via studies that directly compare A with B and B with C) evidence, concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. The only head-to-head RCT comparing modalities found that fluoxetine, alone and in combination with CBT, was superior to pill placebo, though CBT alone was not [@marchFluoxetineCognitiveBehavioralTherapy2004]. Also, the addition of psychotherapy to standard care did not improve outcomes (REF). Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions?

@@

<!-- Indeed, comparing outcomes (Y) between psychotherapy and medication trials requires us to contrast what is called the sample average treatment effect $\tau$ of each, defined in the following way: -->

<!-- $$ -->

<!-- \tau_{psy} =  E(Y_{1} - Y_{0} \,|\, S, psy) -->

<!-- $$ $$ -->

<!-- \tau_{med} =  E(Y_{1} - Y_{0} \,|\, S, med)\  -->

<!-- $$ -->

<!-- where the operator E denotes the expectation over the differences in outcomes between those who received the intervention (T = 1) and those who received the control condition (T = 0), for each sample, $S$, where psy and med stand for psychotherapy and medication respectively. -->

<!-- Obviously, this comparison rests on the assumption that trials in both modalities sample from the same population, $P$, of patients. Formally, this can be expressed as follows: -->

<!-- $$E(Y_{1} \,|\, \text{S, psy}) = E(Y_{1} \,|\, P, \text{psy})$$ -->

<!-- $$E(Y_{1} \,|\, \text{S, med}) = E(Y_{1} \,|\, P, \text{med})$$ -->

<!-- signifying that the effect found in the population would be expected to be found in the same population for each treatment. -->

@@

<!-- In order to compare depression severity across the variety of instruments the studies used, we performed a min-max normalisation to turn each study arm mean score at baseline into a percentage using the following formalism: -->

<!-- $$ -->

<!-- \text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}} -->

<!-- $$ -->

<!-- where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus: -->

<!-- $$ -->

<!-- \text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}} -->

<!-- $$ -->

<!-- where $\text{SD}_{X}$ is the original standard deviation of the mean at baseline. -->

@@

<!-- which we defined following [@lakensCalculatingReportingEffect2013, @cummingUnderstandingNewStatistics2013] as: -->

<!-- $$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$ -->

<!-- where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations.  -->

@@

<!-- If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani et al.'s meta-analysis. -->

For meta-analysis it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation (REF), we simulated n=1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in subsequent analyses (please refer to Supplement for full details).

<!-- This is calculated according to: -->

<!-- $$ SE_{SMD} = \sqrt{\frac{2(1 -r_{t_{1}t_{2}})}{n} + \frac{SMD^2}{2n}}$$ -->

<!-- where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations. -->

<!-- In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics: -->

<!-- $$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$ -->

<!-- for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent meta-analyses. -->

@@

<!--  Unlike the traditional random effects meta-analysis, which assumes that each study's true effect size $\theta_{k}$ varies due to heterogeneity between studies, our multilevel model accounts for the hierarchical structure of the data, with study arms nested within study IDs. This allows us to model variability at both the study level and the study arm level. The multilevel model assumes that each study's true effect size $\theta_{ij}$ (where $i$ indexes the study and $j$ indexes the study arm) is influenced by both the variability between studies and the variability between arms within each study. The model can be expressed as: -->

<!-- $$Yij ∼ \mathcal{N}(\theta_{ij}, \sigma_{ij}^2)$$ -->

<!-- where, -->

<!-- $$θij​∼N(xij​β,τi2​)$$ -->

<!-- where $Y_{ij}$ is the observed effect size for the $j$th arm in the $i$th study, which has a normal distribution with mean $\theta_{ij}$ and sampling error variance $\sigma_{ij}^2$. The true effect size $\theta_{ij}$ is modeled as a study-specific effect with an additional term representing the variability between arms within the study. -->

<!-- This gives rise to the following model: -->

<!-- $$ Yij=xijβ+ui+vij+ϵij $$ -->

<!-- where, -->

<!-- $$ ui​∼N(0,τ2) $$ -->

<!-- {#eq-10} describes the deviation of each study from the overall mean effect size, and -->

<!-- $$vij∼N(0,τi2)$$ -->

<!-- describes the deviation of each arm from the study-specific effect, with $\tau_{i}^2$ representing the heterogeneity within studies. Finally, -->

<!-- $$\epsilon_{ij} \sim \mathcal{N}(0, \sigma_{ij}^2)$$ -->

<!-- {#eq-12} represents the sampling error. -->

<!-- In this framework, we can model the means for each arm of the trials as follows: -->

<!-- $$ \begin{aligned} Υ_{ij} &= \begin{cases} 0 & \text{MedControl:} \quad b_0 + u_i + v_{ij} + \epsilon_{ij}\\ 1 & \text{MedActive:} \quad b_0 + b_{1_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 2 & \text{PsyActive:} \quad b_0 + b_{2_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 3 & \text{PsyControl:} \quad b_0 + b_{3_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ \end{cases} \end{aligned} $$ -->

<!-- {#eq-13} Here, the mean effect size for each level is the sum of $b_0$, the intercept for the reference category (medication control), with the coefficient for each level (e.g., $b_{3_{ij}}$ for psychotherapy controls). The variability between studies and arms within studies is captured by $u_i$ and $v_{ij}$, respectively. The confidence intervals for the means are constructed using the standard errors of the means, which account for the hierarchical structure of the data. Each coefficient represents the contrast between the reference category and each level. For example, $b_{3_{ij}}$ represents the contrast between psychotherapy and medication control arms. Inference on these contrasts is conducted using the following test statistic: -->

<!-- $$ z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})} $$ -->

<!-- {#eq-14} This test allows us to assess the significance of the differences between treatment effects across study arms. -->

@@

```{r, warning=FALSE, message = FALSE, echo = FALSE, output = FALSE }
# my alpha for the z 
target_probability <- 1 - 0.05

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# We may have multiple controls per study id for this reason, though this isn't a problem 
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_multi_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
coef_and_se_means_cbt_fluox_esc_study <- left_join(coef_and_se_means_cbt_fluox_esc_study, study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

@@

@@

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_var2_study <- run_multi_metaregression(list_var2_study, model_1)

@@

# extract SMDs from the list for clinical studies
means_var2_study <-  lapply(list_model_1_var2_study, extract_coefficients_func)

@@

# add the ns to the dataframe
coef_and_se_means_var2_study <- left_join(coef_and_se_means_var2_study, study_count_var2_study[study_count_var2_study$is_missing == FALSE, ] %>% 
                                          ungroup() %>% 
                                          select(condition, n),  
                                        by = "condition")

@@

tbl-baseline-results summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. medication and psychotherapy) and the p-value derives from the test for subgroup differences.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
# looking at CDRS studies only, this is with only 2 levels

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
## this is for four levels
### get CDRS studies
cdrs_studies_for_baseline_metareg <- list_cdrs_study[[1]] # created this above, you can take any of these, as the means and sds are repeated.

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms. As seen in fig-plot-means-all there were substantial differences between the four arms of the meta-analysis with striking differences between the medication and psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ).

@@

list_df_simulated[[1]] %>%
  filter(arm_effect_size == "cohens_d_control") %>%
  ggplot(aes(x = cohens_d, y = 0, color = as.factor(psy_or_med), size = baseline_n)) +
  geom_jitter(height = 0.2, alpha = 0.7) +
  geom_point(alpha = 0.7) +
  labs(
    x = "SMD",
    y = "",
    color = "Type of Study",
    size = "Baseline n"
  ) +
  scale_color_discrete(name = "Type of Study", labels = c("Medication", "Psychological Therapy")) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
### Number of sites
# Write results, tabulate.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

>>>>

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Raphaelle Delpech
    email: raphaelle.delpech.18@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Glyn Lewis
    email: glyn.lewis@ucl.ac.uk
    affiliations:
        - ref: ucl_1        
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: "bibliography/word-style-reference-doc.docx"
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    number-sections: true
execute: 
  echo: true
  auto: true
  warning: FALSE
bibliography: "bibliography/references.bib"
csl: "bibliography/vancouver_style.csl"
font:
  family: Calibri
fig-cap-location: top
---

@@

```{r library, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}

@@

# For referencing
bib_data <- ReadBib(here("multilevel_metaregression", "bibliography", "references.bib"))

```{r global_options, include=FALSE}

knitr::opts_chunk$set(tab.cap.fp_text = officer::fp_text_lite(italic = FALSE, bold = TRUE))
```

@@

For adolescent depression, there are limited head-to-head trials of medication and psychotherapy, and hence recommendations are derived from indirect comparisons of treatment efficacy. The National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend psychotherapy (specifically cognitive behaviour therapy (CBT) or interpersonal therapy (IPT)) over medication in most cases [@niceDepressionChildrenYoung2019]. This is in keeping with two sources of evidence relating to child and adolescent depression: meta-analyses of medication randomised controlled trials (RCTs) that cast doubt on the efficacy of antidepressants, with the exception of fluoxetine [@ciprianiComparativeEfficacyTolerability2016]; and meta-analyses of psychotherapy RCTs that conclude psychotherapy to be efficacious [@cuijpersEffectsPsychologicalTreatments2021]. However, a recent network meta-analysis (NMA) [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments using both direct and indirect (i.e. treatment A with treatment C, via studies that directly compare A with B and B with C) evidence, concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. The only head-to-head RCT comparing modalities found that fluoxetine, alone and in combination with CBT, was superior to pill placebo, though CBT alone was not [@marchFluoxetineCognitiveBehavioralTherapy2004]. Also, the addition of psychotherapy to standard care did not improve outcomes [@goodyer_randomised_2008]. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions?

@@

@@

@@

@@

For meta-analysis it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation [@cuijpers_pre-post_2017], we simulated n=1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in subsequent analyses (please refer to Supplement for full details).

@@

@@

```{r set-up-z, warning=FALSE, message = FALSE, echo = FALSE, output = FALSE }
# my alpha for the z 
target_probability <- 1 - 0.05

@@

```{r define-functions, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

@@

```{r import-dataset, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r metaregression-all, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-fluox, warning=FALSE, message = FALSE, echo=FALSE}
# # B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
# #  grab ids for those studies
# cbt_fluox_esc_study_ids <-
#   df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id
# 
# # Use those ids to subset all dataframes in the list
# # By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# # So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# # This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# # We may have multiple controls per study id for this reason, though this isn't a problem 
# # conceptually.
# 
# list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
#   df %>%
#     filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
#     filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
# })
# 
# # count the number of studies
# cbt_fluox_esc_study <- df_long_for_metan %>%
#   filter(new_study_id %in% cbt_fluox_esc_study_ids,
#          !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
# study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# # CB checked, looks good now.
# 
# # apply the metareg function to the list of cbt, fluox, esc
# model_1 <- as.formula(~ four_level_var)
# list_model_1_cbt_fluox_esc_study <- run_multi_metaregression(list_cbt_fluox_esc_study , model_1)
# 
# # extract coefficients and model characteristics.
# condition <- levels(list_df_simulated[[1]]$four_level_var)
# aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
# aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)
# 
# # extract SMDs from the list for cbt fluox escit
# means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)
# 
# # calculate mean SMDs and ses
# condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
# coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
#  coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))
# 
# # add the ns to the dataframe
# coef_and_se_means_cbt_fluox_esc_study <- left_join(coef_and_se_means_cbt_fluox_esc_study, study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing == FALSE, ] %>% 
#                                         ungroup() %>% 
#                                       select(condition, n),  
#                                     by = "condition")
```

```{r metaregression-nowl, warning=FALSE, message = FALSE, echo=FALSE}

@@

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-clin, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-cdrs, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-hamd, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r metaregression-postsd, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r metaregression-var, warning=FALSE, message = FALSE, echo=FALSE}

@@

# apply the metareg function to the list
model_1 <- as.formula(~ four_level_var)
list_model_1_var2_study <- run_multi_metaregression(list_var2_study, model_1)

@@

# extract SMDs from the list
means_var2_study <-  lapply(list_model_1_var2_study, extract_coefficients_func)

@@

# add the ns to the dataframe
coef_and_se_means_var2_study <- left_join(coef_and_se_means_var2_study, study_count_var2_study[study_count_var2_study$is_missing == FALSE, ] %>%
                                          ungroup() %>%
                                          select(condition, n),
                                        by = "condition")
```

@@

@tbl-baseline-results summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. medication and psychotherapy) and the p-value derives from the test for subgroup differences.

```{r characteristics-metanalysis, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r baseline-severity, echo=FALSE, message=FALSE, warning=FALSE}
# looking at CDRS studies only, this is with only 2 levels

@@

```{r baseline-severity-hamd-cdrs, warning=FALSE, message = FALSE, echo=FALSE}
## this is for four levels
### get CDRS studies
cdrs_studies_for_baseline_metareg <- list_cdrs_study[[1]] # created this above, you can take any of these, as the means and sds are repeated.

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r``round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms. As seen in @fig-plot-means-all there were substantial differences between the four arms of the meta-analysis with striking differences between the medication and psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ).

@@

list_df_simulated[[1]] %>%
  filter(arm_effect_size == "cohens_d_control") %>%
  ggplot(aes(x = cohens_d, y = 0, color = as.factor(psy_or_med), size = baseline_n)) +
  geom_jitter(height = 0.2, alpha = 0.7) +
  geom_point(alpha = 0.7) +
  labs(
    x = "SMD",
    y = "",
    color = "Type of Study",
    size = "Baseline n"
  ) +
  scale_color_discrete(name = "Type of Study", labels = c("Medication", "Psychological Therapy")) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )

 png("plot_means_multi_all.png")
 plot_means_multi_all
 dev.off()
```

@@

```{r baseline-adjust-means, warning=FALSE, message = FALSE, echo=FALSE}

@@

```{r trial-sites, warning=FALSE, message = FALSE, echo=FALSE}
### Number of sites
# Write results, tabulate.

@@

```{r sex-all-studies, warning=FALSE, message = FALSE, echo=FALSE}
